Interactive Summarization of Relational Tables

### 2014-11-14 TableSummarization/Paper/TableSummarization.tex
split, first pass on formal

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{ \textcolor{red}{\bf ||Ins: #1||}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

@@

@@

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints. The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that satisfy that pattern. For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column. We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table. We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and an optimized sampling scheme for efficiently summarizing large tables. Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
\end{abstract}

\section{Introduction}

Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it. Large amounts of data are hard for a human to make sense of. This necessitates summarizing the data in an informative, interpretable way. The summary also needs to be fast and interactive to facilitate deeper exploration.

In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints. A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}. Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a black box `weight function' as user input. In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `weight' in this case denoting the number of non-$\star$ values of the rule for each rule. The weight function can in general be customized by the user.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

% Things to work into example : Number of rules. Ordering.

The first rule tells the user that there are $6000$ tuples in all. We call it an `empty' rule, because it has only $\star$ values. The three additional rules give more information about the table. The number of additional rules ($3$ in this case) is a parameter that can be set by the user. The rules are shown in decreasing order by weight.

The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that Target is selling a lot of bicycles. The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores. The last rule tells the user that walmart is doing well in general over multiple products and states.

This gives the user an overall summary of the results. But after seeing this summary, the user may want to dig deeper to get more details on certain rules. For instance, the user may want to know which states walmart has more sales in, or which products they sell the most. In this case, the user can click on the rule with walmart, and will see an expanded summary as given in table~\ref{table:introexample2}.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}

The three new rules in this table give the user additional information about the $1000$ walmart tuples. This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.

In our summary, we want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible. Thus choosing rules involves a tradeoff between highly specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$). Our decision on what rules to choose is influenced by a rule-weighting scheme, which is further discussed in Section~\ref{sec:weighting}. Briefly, we allow the user to specify a weighting function $W$ as a black box, and display the optimal set of rules as per that weighting function.

This example also raises questions on what operations to allow the user to perform to dig deeper into the rules. We describe our user interface in Section \ref{sec:interface}. 
\end{example}

To summarize, we attempt to summarize relational data in a way that is:
\squishlist
\item \textbf{Flexible : } We allow users to specify a weight function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
\item \textbf{Online : } Does not require any heavy pre-computation. 
\squishend

There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}. These papers mostly focuses on giving the most 'surprising' information to the user i.e. information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution. For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user. In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible. Our approaches are thus complementary and serve different purposes.
% TODO: max-entropy is hard to interpret for human, even though formally very clear. Our thing also works well when user doesn't know the set of column values. The set, at least number of column values, are needed for imagining the max-entropy distribution.

OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'. They have an operation called drill-down, which allows a user to get counts for rules having values in a given column. For instance in our example, a user can first drill down along the `Department Store' column, which will show the total number of tuples for each Department Store. After that, the user may drill down along the product column for `Walmart', which will show the user the number of tuples for each product in Walmart. Thus the user is required to decide which columns to drill down along in order to get interesting or frequent rules, and requires $k$ drill downs to get to rules of size $k$. In contrast, our summary may immediately display a size $k$ rule to the user if it has sufficiently high weight and count. There is no notion of ranking or ordering of rules in OLAP, since all possible rules are displayed for any drill down. In contrast, we only display the top few rules in terms of their score, and display them in decreasing order by weight. This makes the summary shorter and more readable by focusing only on the frequently occurring values in columns.

We further discuss related work in Section~\ref{sec:related}.

\stitle{Overview of paper:} 
\squishlist

\item In Section~\ref{sec:preliminaries}, we introduce the relevant notation and formally define our problem.

\item In Section~\ref{sec:weighting}, we consider different schemes for weighting rules, along with their practical significance.

\item In Section~\ref{sec:interface}, we describe our user interface.

\item In Section~\ref{sec:algorithms}, we present our algorithms for finding optimal rule-based summaries, as well as our sampling schemes for dealing with large tables.

\item In Section~\ref{sec:experiments}, we experimentally evaluate the performance of our algorithms on a real dataset.

\item In Section~\ref{sec:related}, we describe related work.
\squishend

\section{Preliminaries} 
\label{sec:preliminaries}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table.

Our objective (formally defined later) is to summarize the table with a list of {\em rules}. A {\em rule} is a tuple with a value for each column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all columns.

A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. 
%TODO: Give name to out system (and paper)

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e. there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

When a user clicks on a rule to know more about the part of the table covered by that rule, we display a rule-list of sub-rules of the rule.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.

We now define some additional properties of rules. Let $r$ be a rule. Then:
\squishlist
\item {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
\item {\em MCount}($r$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user, to increase the coverage of the rule-list. 
\item {\em Size}($r$) refers to the number of non-starred values in rule $r$. 
\item We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. 
\squishend

The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$ Our problem is to choose the rule-list that maximises total score. We use $MCount$ rather than Count to increase coverage of the table. If we had defined total score as $\sum_{r \in R} \text{Count}(r)W(r)$, then our optimal rule-list would contain rules that repeatedly refer to the most `summarizable' part of the table. For instance, if $a$ and $b$ were the most common values in columns $A$ and $B$, then the summary could consist of rules $(a, b, \star)$, $(a, \star, \star)$, and $(\star, b, \star)$, which tells us nothing about the part of the table with values other than $a$ and $b$.

We now formally define our problem:

\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r)MCount(r)$$
for one of the following interactions:
\begin{enumerate}
\item $[$Rule Interaction$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star Interaction$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{enumerate}
\end{problem}

% Describe architecture here? (With a diagram). Show SampleHandler, QueryEngine, etc.

\section{Weighting Rules}
\label{sec:weighting}
We now describe how we score rules. At a high level, we want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is. Given this weight per rule, our total score for our rule-list is given by : $\sum_{r\in R} MCount(r)W(r)$. Let us consider some forms for the function $W(r)$ and their interpretation:
\begin{enumerate}
\item $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :

Here we set weight equal to the size of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those. Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.

\item $W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column. The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes log($|c|$) bits of information. A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table. Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score. 
\end{enumerate}

Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
\begin{enumerate}
\item Non-negativity : For all rules $r$, $W(r) \geq 0$ 
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\end{enumerate}

\begin{table}
\centering
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\end{table}

\section{Our interface}
\label{sec:interface}
We now explain how the user can interact with our system to explore the contents of a table.

When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}. At any point, the user can click on either a rule, or a star within a rule, to `expand' the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:trivial} would display table~\ref{table:introexample}. Clicking further on the second rule would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Table with only the trivial rule \label{table:trivial}}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$. We can reduce Problem~\ref{prob:optimal-subrule-list} to the following simpler problem by removing the user-interaction based constraints:

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
\end{problem}

$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.

Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\squishlist
\item $[$Rule Interaction$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Interaction$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\squishend

We will describe a greedy algorithm for finding the approximately highest scoring rule-list. First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
\end{lemma}
\begin{proof}
We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$. For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to. Then Score can be rewritten as:
$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
Thus each tuple $t$ `contributes' $W(t)$ to the total score. Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$. Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
as required.
\end{proof}

Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. This gives us 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
\end{problem}

The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.

\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

\begin{lemma}\label{lemma:submodular}
For a given table $T$, the Score function over sets $S$ of rules, defined by :
$$\text{Score}(S) = \sum_{r \in S} MCount(r)W(r)$$
is submodular.
\end{lemma}
\begin{proof}
For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$ 
Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.

Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
which means that Score is submodular.
\end{proof}

Now consider the greedy algorithm:

\begin{framed}
\begin{enumerate}
\item Set $S = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
\item $S = S \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\end{framed}

Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled Algorithm~\ref{algo:best-rule-set}. The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, our algorithm is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$. Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, it's Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set are small.

% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary. Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.

The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty, and then iterates for $k$ steps, adding the best marginal rule at each step. To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.

In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}. We compute the best marginal rule over several passes, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, b, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

We now describe the function to find the best marginal rule. The pseudo-code for the function is in the box titled Algorithm~\ref{algo:best-marginal-rule}. The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far. The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted in all previous iterations. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous iteration. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

For the second pass onwards, we are more selective about which rules to consider for marginal value evaluation. We first set $C_n$ to be the set of rules of size $j$ which are super-rules of rules from $C_o$. Then for each rule $r$ from $C_n$, we consider the known marginal values of its sub-rules from $C$, and use them to upper-bound the marginal value of all super-rules of $r$, as shown in Step 3.3.2. Then we delete from $C_n$ the rules whose marginal value upper bound is less than the currently known best marginal value, since they have no chance of being returned as the best marginal rule. Then we make as actual pass through the table to compute the marginal value of the rules in $C_n$, as shown in Step 3.5. If in any round, the $C_n$ obtained after deleting rules is empty, then we terminate the algorithm and return the highest value rule.

\begin{algorithm}\label{algo:best-rule-set}
\KwIn{$k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)}
\KwOut{$S$ (Solution set of rules)}
$S = \phi$

\For {$i$ from $1$ to $k$}{
$R_m = \text{Find\_best\_marginal\_rule}(S, T, m_w, W)$ \tcc*{Calling Algorithm~\ref{algo:best-marginal-rule}}

$S = S \cup \left\lbrace R_m \right\rbrace$
}
\Return $S$
\caption{Greedy Algorithm for Problem~\ref{prob:optimal-rule-set}}
\end{algorithm}

>>>>

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

@@

\newcommand{\calD}{\mathcal{D}\xspace}

@@

\begin{abstract}
In this paper we present {\em smart drill down},
a system for interactively exploring a relational table
to discover ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a $1000$ tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively drill down on an existing rule to
explore that part of the table.
We describe an
algorithm for finding the approximately optimal list of rules, and an optimized
sampling scheme for efficiently summarizing large tables. Finally, we
demonstrate the usefulness of smart drill down by performing experiments on
a real dataset.
\end{abstract}

\input{intro}

\input{formal}

\input{algorithms}

\input{exp}

\input{extensions}

\input{related}
